### 2018-11-23 paper/paper.tex
Radical rewrite time

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 :

\documentclass[runningheads]{llncs}

\usepackage{complexity}
\usepackage{amsmath}
% \usepackage[ruled,vlined]{algorithm2e}
\usepackage{tikz}
\usepackage{cleveref}

% \usepackage{showframe}

\newcommand{\neighbourhood}{\operatorname{N}}
\newcommand{\vertexset}{\operatorname{V}}

\newcommand{\siplabel}[1]{\label{line:sip:#1}}
\newcommand{\siplineref}[1]{line~\ref{line:sip:#1}}
\newcommand{\siplinerangeref}[2]{lines~\ref{line:sip:#1} to~\ref{line:sip:#2}}

\newcommand{\citet}[2]{#1\cite{#2}}

% cref style
\crefname{figure}{Figure}{Figures}
\Crefname{figure}{Figure}{Figures}
\crefname{section}{Section}{Sections}
\Crefname{section}{Section}{Sections}
\crefname{proposition}{Proposition}{Propositions}
\Crefname{proposition}{Proposition}{Propositions}
\crefname{corollary}{Corollary}{Corollaries}
\Crefname{corollary}{Corollary}{Corollaries}

% \crefname{algocf}{Algorithm}{Algorithms}
% \Crefname{algocf}{Algorithm}{Algorithms}

\usetikzlibrary{decorations, decorations.pathreplacing, decorations.pathmorphing,
    calc, backgrounds, positioning, tikzmark, patterns, fit}

\definecolor{highlight}{rgb}{0.529412, 0.74902, 0.466667}

\begin{document}

\title{Sequential and Parallel Solution-Biased Search for Subgraph Algorithms\thanks{This work was
supported by the Engineering and Physical Sciences Research Council (grant numbers EP/P026842/1,
EP/M508056/1, and EP/N007565). This work used the Cirrus UK National Tier-2 HPC Service at EPCC
(http://www.cirrus.ac.uk) funded by the University of Edinburgh and EPSRC (EP/P020267/1).}}

\author{Blair Archibald\inst{1}\orcidID{0000-0003-3699-6658}\and
Fraser Dunlop\inst{2}\orcidID{0000-0002-4485-4871}\and
Ruth Hoffmann\inst{2}\orcidID{0000-0002-1011-5894}\and
Ciaran McCreesh\inst{1}\orcidID{0000-0002-6106-4871}\and
Patrick Prosser\inst{1}\orcidID{0000-0003-4460-6912} \and
James Trimble\inst{1}\orcidID{0000-0001-7282-8745}}

\authorrunning{B. Archibald et al.}

\institute{University of Glasgow, Glasgow, Scotland \and University of St Andrews, St Andrews,
Scotland \\ \email{ciaran.mccreesh@glasgow.ac.uk}}

\maketitle

\begin{abstract}
    The current state of the art in subgraph isomorphism solving involves using degree as a
    value-ordering heuristic to direct backtracking search. Such a search makes a heavy commitment
    to the first branching choice, which is often incorrect. To mitigate this, we introduce and
    evaluate a new approach, which we call ``solution-biased search''. By combining a
    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we achieve two
    orders of magnitude faster performance on satisfiable instances, whilst not affecting
    performance on unsatisfiable instances. This search can be parallelised in a very simple
    but effective way, giving us a further speedup of over ?? from thirty-six cores, across both
    satisfiable and unsatisfiable instances.  We also show that solution-biased search is also
    suitable for optimisation problems, by using it to improve two maximum common induced subgraph
    algorithms.
\end{abstract}

\section{Introduction}

The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
inside a larger ``target'' graph. Although the problem is \NP-complete, by combining design
techniques from artificial intelligence with careful algorithm engineering, modern subgraph
isomorphism solvers can often produce exact solutions quickly even on graphs with thousands of
vertices. The current single strongest subgraph isomorphism solver uses ``highest degree first'' as
a value-ordering heuristic to direct a constraint programming style search
\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:journals/jair/McCreeshPST18}. This
heuristic is much better than branching randomly, but is still far from perfect. To offset mistakes
made by this heuristic, this paper proposes a new perspective on value-ordering: rather than
defining a search order, we use degree to direct what \emph{proportion} of the search effort should
be spent in each subproblem. By combining rapid restarts and nogood recording, and introducing a
small amount of randomness into the value-ordering heuristic, we make a state-of-the-art subgraph
algorithm perform two orders of magnitude better on a large number of satisfiable instances, whilst
performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.

This new form of search can also be parallelised. By running many threads with different random
seeds but the same restart schedule, and sharing nogoods only following restarts, we can achieve
speedups of ??, with a \emph{much} simpler implementation than conventional work-stealing.  Finally,
we show that this strategy is also effective in an optimisation setting, producing benefits in two
maximum common induced subgraph algorithms. This combination of techniques gives us, for the first
time, a practical alternative to backtracking search for constraint programming style algorithms
where we have a strong value-ordering heuristic and where we care both about satisfiable and
unsatisfiable instances.

\subsection{Definitions}

% Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
% that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself (a \emph{loop}). The
% \emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
% the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood. When the graph is clear from the context, we omit
% the $G$ subscripts. As per \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
% vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
% simple paths of length $\ell$ between vertices $v$ and $w$ in $G$.

The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from the
vertices of a \emph{pattern} graph $\mathcal{P}$ to a \emph{target} graph $\mathcal{T}$, such that
adjacent vertices in $\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$ (including that
vertices with loops in $\mathcal{P}$ may only be mapped to vertices with loops in $\mathcal{T}$).
The \emph{induced} problem additionally requires that non-adjacent vertices are mapped to
non-adjacent vertices.

?? Applications

% We write $v \mapsto v'$ to mean that pattern vertex $v$ is mapped to target $v'$ under
% either kind of mapping. Finally, we define a function $\FuncSty{compatible}(v, w)$ to mean that
% mapping $v \mapsto w$ is feasible based upon both the loop rule, and the neighbourhood degree
% sequence relation of \citet{Zampelli et al.\ }{DBLP:journals/constraints/ZampelliDS10}.

\subsection{The Glasgow Subgraph Solver}

This paper looks at improving the Glasgow Subgraph
Solver\footnote{\url{https://github.com/ciaranm/glasgow-subgraph-solver/}}, which can solve both the
non-induced and the induced subgraph isomorphism problems.  The solver is very closely based upon
the $k{\downarrow}$ algorithm of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} with $k =
0$, and we refer the reader to that paper for full technical details; that algorithm, in turn, is a
simplification and re-engineering of an older Glasgow algorithm
\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.  Essentially, the solver is essentially
a dedicated forward-checking constraint programming implementation specifically for subgraph
problems. It works with a model having a variable per pattern graph vertex, with domains ranging
over the target graph vertices, and performs a backtracking search to map each pattern vertex to a
target vertex whilst propagating adjacency and injectivity constraints (together with further
implied constraints based upon degrees and paths).  However, it uses specialised bit-parallel data
structures and algorithms, and propagates constraints in a fixed order rather than using a queue.

% We now briefly outline the non-shaded parts of \cref{algorithm:sip}, which is how we will solve the
% subgraph isomorphism problem. This algorithm is very closely based upon the $k{\downarrow}$
% algorithm of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
% for full technical details; that algorithm, in turn, is a simplification of the Glasgow2 algorithm
% \cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}. We show the non-induced problem; the
% induced problem may be solved by including the loop-complements of $\mathcal{P}$ and $\mathcal{T}$
% in $L$ on \siplineref{supplemental}.
% 
% \Cref{algorithm:sip} works by recursively building up a valid mapping of pattern vertices to target
% vertices. For each pattern vertex $v$, we have a domain $D_v$ containing a set of feasible target
% vertices. Each domain is initialised at the top of search (\siplinerangeref{initstart}{initend}),
% using the invariants described by \citet{Zampelli et al.\ }{DBLP:journals/constraints/ZampelliDS10} and
% \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17} to restrict its initial values.
% 
% The $\FuncSty{search}$ procedure recursively selects an unassigned domain $D$
% (\siplineref{variableordering}), and tries assigning
% it each of its remaining values in turn (\siplineref{valueordering}). The effects of this assignment
% are propagated (\siplineref{dopropagate}). If an
% inconsistency is detected, we continue and try another value; otherwise, we recurse
% (\siplineref{recurse}) until a solution is found. Finally, if we run out of values, we backtrack
% (\siplineref{backtrack}).
% 
% Our $\FuncSty{propagate}$ routine is also close to that of \citet{Hoffmann et al.\ }{DBLP:conf/aaai/HoffmannMR17},
% differing only in that we carry out the filtering for injectivity
% (\siplinerangeref{alldiffstart}{alldiffend}) after every assignment rather than once at the end of
% the unit propagation loop. (We make this change because the all-different propagator used is not
% idempotent, and we wish to avoid an odd interaction when we introduce nogood filtering.)

% \begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
%     \begin{tikzpicture}[remember picture,overlay]
%         \coordinate (restart1sc) at ($(pic cs:restart1s) + (0, 0.15)$);
%         \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
%         \node [fill=highlight, rounded corners, fit=(restart1sc) (restart1ec)] { };
%         \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
%         \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
%         \node [fill=highlight, rounded corners, fit=(restart2sc) (restart2ec)] { };
%         \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
%         \coordinate (restart9ec) at ($(pic cs:restart9e) + (-0.15, 0.00)$);
%         \node [fill=highlight, rounded corners, fit=(restart9sc) (restart9ec)] { };
%     \end{tikzpicture}
% 
%     \nl $\FuncSty{subgraphIsomorphism}$ (Graph $\mathcal{P}$, Graph $\mathcal{T}$) $\rightarrow$ Bool \;
% \nl \Begin{
%     \nl \lIf{$\left|\vertexset(\mathcal{P})\right| >
%     \left|\vertexset(\mathcal{T})\right|$}{$\KwSty{return}~\KwSty{false}$\siplabel{enough}}
%     \nl Discard isolated vertices in $\mathcal{P}$\siplabel{isolated} \;
%     \nl \KwSty{global} $L \gets \big[ (\mathcal{P},\ \mathcal{T}), \siplabel{supplemental}
%         (\mathcal{P}^{1,2}, \mathcal{T}^{1,2}),
%         (\mathcal{P}^{2,2},\ \mathcal{T}^{2,2}),$ \\
%         $\hspace*{3em}(\mathcal{P}^{3,2},\ \mathcal{T}^{3,2}),
%         (\mathcal{P}^{4,2},\ \mathcal{T}^{4,2}) \big]$ \;
%     \nl \ForEach{\siplabel{initstart}$v \in \vertexset(\mathcal{P})$}{
%         \nl $D_v \gets \vertexset(\mathcal{T})$\;
%         \nl \ForEach{$(P,\,T) \in L$}{\nl$D_v \gets \{ w \in D_v : \FuncSty{compatible}(v, w)\}$\siplabel{initend}
%         }
%     }
%     \nl \While{$\KwSty{true}$\siplabel{restartsloop}}{
%         \nl \lIf{$\KwSty{not}~\FuncSty{propagate}(D)$}{$\KwSty{return}~\KwSty{false}$}
%         \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0, \textnormal{the next Luby sequence value})$\tikzmark{restart1e} \;
%         \nl $S \gets \FuncSty{search}(\{ E \in D : \left|E\right| > 1 \}, \top)$ \;
%         \nl \tikzmark{restart9s}\lIf{$S \ne \KwSty{restart}$}{\tikzmark{restart9e}$\KwSty{return}~S$}
%         \nl \tikzmark{restart2s}trigger nogoods containing only one clause\tikzmark{restart2e} \;
%     }
% }
% \caption{The main subgraph isomorphism algorithm. The shaded code is added for restarting. See
%     \cref{algorithm:sip:search,algorithm:sip:propagate} for the search and propagation routines.}
% \label{algorithm:sip}
% \end{algorithm}
% 
% \begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
%     \begin{tikzpicture}[remember picture,overlay]
%         \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
%         \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
%         \node [fill=highlight, rounded corners, fit=(restart3sc) (restart3ec)] { };
%         \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
%         \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
%         \node [fill=highlight, rounded corners, fit=(restart4sc) (restart4ec)] { };
%         \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
%         \coordinate (restart5ec) at ($(pic cs:restart5e) - (0, 0.10)$);
%         \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
%         \node [fill=highlight, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
%         \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
%         \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
%         \coordinate (restart6rc) at ($(pic cs:restart6r) + (0.8, 0.00)$);
%         \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec) (restart6rc)] { };
%         \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
%         \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
%         \node [fill=highlight, rounded corners, fit=(restart8sc) (restart8ec)] { };
%     \end{tikzpicture}
%     \nl $\FuncSty{search}$ (Domains $D$, Trail $B$) $\rightarrow$ Bool
%     \tikzmark{restart8s}\KwSty{or} \KwSty{restart}\tikzmark{restart8e} \;
% \nl \Begin{
%     \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
%     \nl $D_v \gets \textnormal{a domain from}~D~\textnormal{chosen by variable heuristic}$\siplabel{variableordering} \;
%     \nl \tikzmark{restart3s}$C \gets \emptyset$\tikzmark{restart3e} \;
%     \nl \ForEach {$v' \in D_v~\textnormal{ordered by value heuristic}$\siplabel{valueordering}}{
%         \nl \tikzmark{restart4s}$C \gets C \cup \{ v' \}$\tikzmark{restart4e} \;
%         \nl $D' \gets \FuncSty{clone}(D)$ \;
%         \nl $D'_v \gets \{ v' \} $ \;
%         \nl \If{\FuncSty{propagate}(D')\siplabel{dopropagate}}{
%             \nl $S \gets \FuncSty{search}(\{ E \in D'\,{:}\,\left|E\right|{>}\,1 \}, B \wedge (v\,{\mapsto}\,v'))$\siplabel{recurse}\;
%             \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
%             \nl \tikzmark{restart5s}\ElseIf{$S = \KwSty{restart}$}{
%                 \nl \ForEach{$c \in C$}{\nl$\textnormal{post nogood}~B \wedge (v \mapsto c) \Rightarrow \bot$\tikzmark{restart5r}}
%                 \nl $\KwSty{return}~\KwSty{restart}$\tikzmark{restart5e} \;
%             }
%         }
%     }
%     \nl \tikzmark{restart6s}\If{$(R \gets R + 1) = R'$\tikzmark{restart6r}}{
%         \nl post nogood $B$ \;
%         \nl $\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$\siplabel{countrestarts}}
%     \nl $\KwSty{return}~\KwSty{false}$\siplabel{backtrack}
% }
% \caption{Search routine for \cref{algorithm:sip}.}
% \label{algorithm:sip:search}
% \end{algorithm}
% 
% \begin{algorithm}[tb]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
%     \begin{tikzpicture}[remember picture,overlay]
%         \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
%         \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
%         \node [fill=highlight, rounded corners, fit=(restart7sc) (restart7ec)] { };
%     \end{tikzpicture}
% \nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
% \nl \Begin{
%     \nl \While{$D_v \gets $\textnormal{a unit domain from}~$D$}{
%         \nl $v' \gets \textnormal{the single value in}~D_v$ \;
%         \nl \tikzmark{restart7s}trigger nogoods with watch $(v \mapsto v')$\tikzmark{restart7e}\siplabel{2wl} \;
%         \nl \ForEach{$D_w \in D \setminus \{ D_v$ \}}{
%             \nl $D_w \gets D_w \setminus \{ v' \}$ \siplabel{removev} \;
%             \nl \ForEach{$(P,\,T) \in L~\KwSty{where}~v \sim_P w$}{\nl$D_w \gets D_w \cap \neighbourhood_T(v')$}
%             \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
%         }
% \nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$\siplabel{alldiffstart} \;
% \nl \ForEach{$D_v \in D$ \textnormal{in ascending cardinality order\siplabel{eachdomain}}}{
%     \nl $(D_v, A,\,n) \gets (D_v \gets D_v \setminus H\siplabel{elimhall}, A \cup D_v,\,n + 1)$ \siplabel{acc} \;
%     \nl \lIf{$D_v = \emptyset~\vee~|A| < n$}{$\KwSty{return}~\KwSty{false}$\siplabel{failhall}}
%     \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}\siplabel{alldiffend}
% }
%     }
% \nl $\KwSty{return}~\KwSty{true}$ \;
% }
% \caption{Propagation routine for \cref{algorithm:sip}.}
% \label{algorithm:sip:propagate}
% \end{algorithm}

% For the variable-ordering heuristic on \siplineref{variableordering} of
% \cref{algorithm:sip}, we use smallest domain first, tiebreaking on highest degree
% \cite{DBLP:journals/ai/HaralickE80,DBLP:journals/jair/McCreeshPST18}. Variable ordering is not the
% focus of this paper---as far as possible, we use this same dynamic ordering throughout the paper, so
% as not to affect the performance of the algorithm on unsatisfiable instances.  For value ordering,
% which occurs on \siplineref{valueordering} of \cref{algorithm:sip}, we will initially select
% vertices from highest degree to lowest degree; value ordering is the major theme of the remainder of
% this paper. Note that the variable-ordering heuristic affects both unsatisfiable and satisfiable
% instances, whilst the value-ordering heuristic is relevant only for satisfiable instances.

\subsection{Experimental Setup}

Our experiments are performed on the EPCC Cirrus HPC facility, on systems with dual Intel Xeon
E5-2695 v4 CPUs and 256GBytes RAM, running Centos 7.3.1611. We use GCC 7.2.0 as the compiler. For
paralellism, we use C++ native threads, and for distributed parallelism we also use the SGI MPT
implementation of MPI. All timing measurements are steady-clock, and we use a deterministic
pseudo-random number generator for reproducibility.

We use the dataset introduced by \citet{Kotthoff et al.\ }{DBLP:conf/lion/KotthoffMS16} for
evaluation. This dataset brings together a range of randomly-generated and application instance
families from earlier papers:

\begin{description}
    \item[BVG(r), M4D(r), and Rand] are families of randomly generated graphs using different models (bounded
        degree, regular mesh, and uniform), where each pattern is a permuted random connected subgraph
        of the target (and so each instance is satisfiable) \cite{DBLP:journals/pami/CordellaFSV04}.
        These benchmark instances are widely used, but have unusual properties and so broad
        conclusions should not be drawn based solely upon behaviour on these instances
        \cite{DBLP:journals/jair/McCreeshPST18}.
    \item[SF] contains randomly generated scale-free graphs using a similar method
        \cite{DBLP:journals/constraints/ZampelliDS10}.
   \item[LV] consists of various kinds of graph gathered by Larrosa and Valiente
       \cite{DBLP:journals/mscs/LarrosaV02} from the Stanford Graph Database. We include both the
        50 small graphs, and the 50 larger graphs.
    \item[Phase] contains hard crafted instances that lie near the satisfiable / unsatisfiable phase
        transition \cite{DBLP:journals/jair/McCreeshPST18}.
    \item[PR] contains graphs generated from segmented images, corresponding to a computer vision
        problem \cite{DBLP:journals/pr/SolnonDHJ15}.
    \item[Images and Meshes] contain graphs representing 2D segmented images and 3D object models
        respectively, again representing a computer vision problem
        \cite{DBLP:journals/cviu/DamiandSHJS11}.
\end{description}

\noindent
Other studies use a random selection of 200 of each of the instances from the ``meshes'' and
``images'' families because some earlier solvers find many of these instances extremely hard. We
would like to have a larger number of satisfiable instances in our test set, and so we include all
pattern / target pairs. This gives a total of 14,621 instances (rather than the original 5,725). At
least \input{gen-n-sat.tex}\unskip\ of these instances are known to be satisfiable for the
non-induced problem, and at least \input{gen-n-unsat.tex}\unskip\ are unsatisfiable.

\section{Improving Sequential Search}

\begin{figure}[tb]
    \includegraphics*{gen-graph-others.pdf}
    \hfill
    \includegraphics*{gen-graph-others-induced.pdf}

    \bigskip

    \includegraphics*{gen-graph-others-zoom.pdf}
    \hfill
    \includegraphics*{gen-graph-others-induced-zoom.pdf}

    \caption{On the top row, the cumulative number of instances solved over time, comparing our algorithm
    (both in its basic form, and with the improvements introduced in the remainder of the paper) to
    other solvers, for the non-induced and induced problems. On the bottom row, the same,
    considering only satisfiable instances.}
    \label{figure:others}
\end{figure}

We begin with a set of baseline performance measurements. In the first and
third plots \cref{figure:others} we show the cumulative number of instances solved over time for the
non-induced and induced problems respectively. We compare the Glasgow Subgraph Solver using
depth-first search (DFS) and with the modifications described in the remainder of this paper
(solution-biased search, SBS), the PathLAD variation of the LAD algorithm
\cite{DBLP:journals/ai/Solnon10,DBLP:conf/lion/KotthoffMS16}, VF2
\cite{DBLP:journals/pami/CordellaFSV04}, RI \cite{DBLP:journals/bmcbi/BonniciGPSF13}, and VF3
\cite{DBLP:conf/gbrpr/CarlettiFSV17} (which only supports the induced problem), in each case using
the original implementation provided by the algorithm's authors. For each algorithm, the $y$ value
gives the cumulative number of instances which (individually) can be solved in no more than $x$
milliseconds.  The vertical distance between two lines therefore shows how many more instances can
be solved by one solver than another, if every instance is run separately with the chosen $x$
timeout. The horizontal distance shows how many times longer the per-instance timeout would need to
be to allow the rightmost algorithm to succeed on $y$ out of the 14,621 instances (bearing in mind
that the two sets of $y$ instances could be different).  The plots show that our starting point
comfortably beats PathLAD, VF2, VF3 and RI, except for very low choices of timeout.

The dataset includes many instances which are extremely easy for a good solver, and so it can be
hard to see the differences between the stronger solvers at higher runtimes. This paper focusses
upon improving the performance on the remaining hard satisfiable instances, and so in the second and
fourth plots in \cref{figure:others} (and in subsequent cumulative plots) we show only satisfiable
instances, and use a reduced range on both axes.  For the remainder of this paper, we show only
the non-induced problem, which tends to be harder; results with the induced variant are similar.

\subsection{Discrepancy Searches}

A \emph{discrepancy} is where search goes against a value-ordering heuristic.  Discrepancy searches
\cite{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07}
are alternatives to backtracking search that initially search disallowing all discrepancies, and
then retry search allowing an increasing number of discrepancies at each iteration until either a
solution is found or unsatisfiability is proven. These schemes assume that value-ordering heuristics
are usually reliable, and that most solutions can be found with only a small number of
discrepancies; in such cases, the heavy commitment to early branching choices made by conventional
backtracking search can be extremely costly.

\begin{figure}[tb]
    \includegraphics*{gen-graph-dds.pdf}
    \hfill
    \includegraphics*{gen-graph-scatter-dds.pdf}

    \caption{Comparing depth-bounded discrepancy search (DDS) to depth-first backtracking search
    (DFS), both using degree as the value-ordering heuristic.}\label{figure:scatter-dds}
\end{figure}

\Cref{figure:scatter-dds} shows the effects of adding Walsh's \cite{DBLP:conf/ijcai/Walsh97}
depth-bounded discrepancy search (DDS) to the solver (results with other discrepancy search variants
are similar). On the scatter plot, each point represents the solving time for one instance---to
avoid noise for easier instances, we measure only time spent during search, and exclude time spent
in preprocessing and initialisation.  Points below the $x-y$ diagonal are speedups, whilst points on
the top and right axes represent instances which timed out after one thousand seconds with one
algorithm, but not the other. For satisfiable instances, the different point styles show the
different families, whilst all unsatisfiable instances are shown as dark dots.  The points well
below the diagonal line and along the right-hand axis on the scatter plot show that DDS can
sometimes be extremely beneficial on satisfiable instances.  However, on both unsatisfiable and most
satisfiable instances, the overheads can be extremely large, and DDS is much worse in aggregate and
is not a viable approach (even when only considering satisfiable instances).  These large overheads
are to be expected: discrepancy searches are aimed primarily at getting better feasible solutions in
optimisation problems which are too large for a proof of optimality to be a realistic prospect, and
they are not well-suited for unsatisfiable decision problems. Despite this, the extremely large
gains on some satisfiable instances confirm our suspicions that we should find an alternative to
heavy-commitment backtracking search.

\subsection{Value-Ordering Heuristics}

\begin{figure}[tb]
    \includegraphics*{gen-graph-value-ordering-heuristics.pdf}
    \hfill
    \includegraphics*{gen-graph-scatter-heuristics.pdf}

    \caption{On the left, the cumulative number of satisfiable instances solved over time, using
    four different value-ordering heuristics. On the right, an instance by instance comparison of
    the degree and biased heuristics on all instances. Points on the outer axes are timeouts, and
    point style shows instance family.}
    \label{figure:value-ordering-heuristics}
\end{figure}

Traditionally, value-ordering heuristics are designed to drive search towards the most promising
region of the search space ?? cite, or the most constrained ?? cite, or the region with the highest solution
density ?? cite. In subgraph isomorphism, this is done by selecting vertices from highest degree to
lowest \cite{DBLP:journals/jair/McCreeshPST18}. The left-hand plot of
\cref{figure:value-ordering-heuristics} demonstrates that this is indeed a good choice: the
\textbf{Degree} heuristic's line shows much better performance on satisfiable instances than the
\textbf{Random} (branch randomly) or \textbf{Anti} (branch from lowest degree to highest) heuristic
lines. (On unsatisfiable instances, the value-ordering heuristic has no effect on performance,
because a complete search must be performed.)

But what happens if our value-ordering heuristic has to choose between mapping a pattern vertex to,
for example, three vertices of degree ten, two vertices of degree nine, or five of degree two? When
driving conventional backtracking search, the degree heuristic would pick one of the vertices of
degree ten, and we would commit all of our search effort to the exponentially large search tree
underneath it.  We will show that this is not a wise choice, and that instead, we should \emph{commit
equal search effort} to each of the three subproblems found by mapping to vertices of degree ten.
And similarly, should we be certain that vertex of degree ten is so much better than a vertex of
degree nine that we should commit no effort to degree nine vertices until all the degree ten
subproblems have been explored? Or might it be better to commit, say, twice as much effort to each
degree ten subproblems as to each degree nine subproblems, and only a very small amount of effort to
the degree two subproblems?  To test this hypothesis, we will now introduce a new alternative to
backtracking search, which we call \emph{solution-biased search}.  This search is made up of three
components: a new slightly-random value-ordering heuristic, rapid restarts, and nogood recording.

\subsection{Biased Value-Ordering}

We define a new \textbf{Biased} value-ordering heuristic, as follows. When branching, we select a
vertex $v'$ from the chosen domain $D_v$ with probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in
D_v}{2^{\deg(w)}}} \text{.} \] This heuristic is now equally likely to pick between vertices of
equal degree, is twice as likely to pick a vertex of degree $d$ as one of degree $d - 1$, and is
over a thousand times more likely to pick a vertex of degree $d$ than degree $d - 10$.
\Cref{figure:value-ordering-heuristics} shows that the performance of the degree and biased
heuristics are very similar: the biased heuristic has slightly worse performance below a ??
timeout, and effectively equal performance above it. This shows that we can introduce an element of
randomness into the degree value-ordering heuristic without adversely affecting its performance
\emph{in aggregate}.

The right-hand plot of \cref{figure:value-ordering-heuristics} gives a detailed comparison of the
degree and biased heuristics. It shows that despite the aggregate performance being similar, on a
case by case basis, the two heuristics can make a large difference to the performance of individual
satisfiable instances. This suggests that although degree is a good heuristic, we should perhaps not
commit heavily to a vertex of highest degree, but also consider vertices of the same or similar
degree.

\subsection{Restarting Search and Nogood Recording}

Having introduced a new value-ordering heuristic, we must now also move away from depth-first
backtracking search. We do this by using \emph{restarts} and \emph{nogood recording}. The general
idea is to perform a certain amount of search, and then if no solution has been found (and
unsatisfiability has not been proven), to abandon search and restart from the beginning. Such an
approach can only be beneficial if something changes after restarting---in a constraint programming
setting, this is usually the variable-ordering heuristic
\cite{DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
In this paper, we instead rely upon randomness in our new \emph{value}-ordering heuristic, and
continue to use smallest domain first with static tiebreaking for variable-ordering.\footnote{It may
be possible to further improve the solver by also introducing randomness or some form of learning
into its variable-ordering heuristic. However, simultaneously introducing a second change would
considerably complicate the empirical analysis. Additionally, the solver's current hand-crafted
variable-ordering heuristics already beat adaptive heuristics like impact or activity-based search.}
Using restarts on value-ordering heuristics is uncommon (although \citet{Razgon et al.\
}{DBLP:conf/flairs/RazgonOP07} look at learning value-ordering heuristics from restarts, \citet{Chu
et al.\ }{DBLP:conf/cp/ChuSS09} use a similar scheme in the context of parallel search, and an early
approach by \citet{Gomes et al.\ }{DBLP:conf/aaai/GomesSK98} does so in an optimisation context).

% The shaded code shows how to introduce restarts to \cref{algorithm:sip}. The recursive
% $\FuncSty{search}$ procedure is modified to return either true (if the instance is satisfiable),
% false (unsatisfiable), or a third value if some threshold was reached without the solution being
% known.  The main function then calls $\FuncSty{search}$ in a loop (\siplineref{restartsloop}), until
% a solution is found.

Preliminary experiments directed us to use the Luby scheme \cite{DBLP:journals/ipl/LubySZ93} to
determine when to restart.
%We count the number of backtracks
%(that is, when we reach the end of the main $\KwSty{foreach}$ loop on
%\siplineref{countrestarts})
%to decide when to restart, rather than counting recursive calls.
Following convention, we multiply each item in the Luby sequence by a constant---we used the SMAC
automatic parameter tuner \cite{DBLP:conf/lion/HutterHL11} to select the value 660.
% The remainder of the changes are to introduce nogoods.  The $B$ argument to $\FuncSty{search}$ tracks the branching choices
% made so far, in the form of a CNF clause with each literal being a guessed $(v \mapsto v')$
% assignment.

To avoid exploring portions of the search
space that we have already visited, every time we restart, we add new constraints to the problem
which eliminate already-explored subtrees---such a constraint is called a nogood. We generate
simple decision nogoods. That is, upon backtracking due to a decision to restart, we post a nogood
of the form $(v \mapsto v') \wedge (w \mapsto w') \wedge (x \mapsto x') \Rightarrow \bot$ for every
branch to the left of the current (incomplete) branch in the search tree, and when we first make a
decision to restart before backtracking, we post a similar nogood eliminating the entire subtree
explored. We use the two watched literals technique \cite{DBLP:conf/dac/MoskewiczMZZM01} to
propagate stored nogoods.
%(\siplineref{2wl}).
This has two benefits: the propagation complexity does not particularly depend upon the number of
stored nogoods, and it does not require any work upon backtracking.  Other more sophisticated nogood
generation and propagation schemes exist \cite{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17},
but these are not helpful in this setting (our solver does not maintain arc consistency or use a
propagation queue).

\subsection{Solution-Biased Search in Practice}

\begin{figure}[tb]
    \includegraphics*{gen-graph-restarts.pdf}
    \hfill
    \includegraphics*{gen-graph-scatter-by-family.pdf}

    \caption{On the left, the number of satisfiable instances solved over time, comparing
    solution-biased search (SBS), random search with restarts (RSR), and the three value-ordering
    heuristics with conventional depth-first search. To the right, a comparison between the original
    algorithm and solution-biased search.}
    \label{figure:old-vs-new}
\end{figure}

In \cref{figure:old-vs-new} we show the effects of adding restarts and nogood recording to the
algorithm. With restarts and nogood recording (random search with restarts, RSR), the random
value-ordering heuristic comfortably beats the degree strategy with depth-first search. In other
words, although having a good value-ordering heuristic is beneficial, introducing randomness into
the search is better, if it is done alongside a mechanism to avoid heavy commitment to any
particular random choice. However, the biased heuristic together
with restarts (solution-biased search, SBS) is better still---that is, if we are introducing
restarts, then it is better to add a small amount of randomness to a tailored heuristic than it is
to throw away the heuristic altogether.  Indeed, the original algorithm can solve ?? satisfiable
instances by ?? seconds, whilst the biased and random restarting algorithms require only ?? seconds
and ?? seconds respectively to solve the same number.

In the more detailed view in the right-hand plot of \cref{figure:old-vs-new}, comparing the original
algorithm to solution-biased search, all of the unsatisfiable instances are very close to the $x-y$
diagonal, showing that their performance is nearly unchanged. On the other hand, there are large
numbers of satisfiable instances well below the diagonal line, indicating large speedups.  Better
yet, there are only a handful of satisfiable instances that are more than a factor of ten times
worse.  In other words, as well as improving performance, we have made up most of the consistency we
lost by introducing randomness.  This is a promising result: for the first time, we have introduced
a viable alternative to backtracking search in a setting where we both have a strong
\emph{value}-ordering heuristic, \emph{and} care about both satisfiable and unsatisfiable instances.

\begin{figure}[tb]
    \includegraphics*{gen-graph-scatter-biased-goods.pdf}
    \hfill
    \includegraphics*{gen-graph-scatter-random.pdf}
    \caption{On the left, not using nogood recording introduces slowdowns, particularly on
    unsatisfiable instances. On the right, using a random value-ordering gives much worse
    performance on many satisfiable instances.}
    \label{figure:features}
\end{figure}

As we might expect, these properties do not hold if any of the combination of changes are disabled.
In the left-hand plot of \cref{figure:features}, we see large slowdowns on unsatisfiable instances
when disabling nogood recording, and on the right-hand plot we see many more satisfiable instances
above the $x-y$ diagonal when using the random value-ordering heuristic as opposed to the
degree-biased heuristic.

\subsection{Why ``Solution-Biased'', and why Exponential Bias?}

?? Although we have shown that it provides good results, we have yet to justify where the biased
formula comes from, or indeed why we call this approach ``solution-biased''. The crux of the
matter is that whilst we do not know where solutions actually lie in the search space, we \emph{do}
know that subproblems under a target vertex of degree $d$ tend to contain a small constant times
more search nodes than those under a target vertex of degree $d - 1$.

\section{Parallel Search}

?? No sharing, sharing with Luby restarts, sharing with constant restarts, sharing with triggered
restarts

?? Reproducibility, constant and triggered

?? How many more nodes on unsat instances

?? Scalability to MPI

\begin{figure}[tb]
    \centering
    \includegraphics*{gen-graph-parallel.pdf}

    \medskip

    \includegraphics*{gen-graph-scatter-par.pdf}
    \hfill
    \includegraphics*{gen-graph-scatter-parconsttick.pdf}

    \caption{Parallel}\label{figure:parallel}
\end{figure}

% \begin{figure}[tb]
%     \includegraphics*{gen-graph-scatter-par-repeat.pdf}
%     \hfill
%     \includegraphics*{gen-graph-scatter-parconsttick-repeat.pdf}
% 
%     \includegraphics*{gen-graph-scatter-parconst-nodes.pdf}
%     \hfill
%     \includegraphics*{gen-graph-scatter-parconsttick-nodes.pdf}
% 
%     \caption{Parallel nodes}\label{figure:parallelnodes}
% \end{figure}

\section{Maximum Common Subgraph Algorithms}

\begin{figure}[tb]
    \centering
    \includegraphics*{gen-graph-mcs.pdf}

    \medskip

    \includegraphics*{gen-graph-scatter-kdown.pdf}
    \hfill
    \includegraphics*{gen-graph-scatter-mcsplit.pdf}

    \caption{Above, the cumulative number of maximum common induced subgraph instances solved by
    k${\downarrow}$ and McSplit${\downarrow}$ over time, with and without restarts and nogood
    recording. Below, comparing k${\downarrow}$ (left) and McSplit${\downarrow}$ (right) on an instance by instance
    basis.}\label{figure:mcs}
\end{figure}

Having looked at subgraph isomorphism in detail, we now briefly discuss the maximum common induced
subgraph problem, to see whether our new approach to search has more general applicability.  Two
recent algorithms for this problem also make use of backtracking search with degree as a
value-ordering heuristic. The k${\downarrow}$ algorithm \cite{DBLP:conf/aaai/HoffmannMR17} attempts
to solve the problem by first trying to solve the induced subgraph isomorphism problem, and then if
that fails, retries allowing a single unmatched vertex (and thus using weaker invariants), and so
on. Due to its similarity to the Glasgow Subgraph Solver, we can introduce the same bias and restart
strategy.

Meanwhile, the McSplit${\downarrow}$ algorithm
\cite{DBLP:conf/ijcai/McCreeshPT17} uses a constraint programming style search, but with special
propagators and backtrackable data structures that exploit special properties of the problem. The
unconventional domain store used by McSplit${\downarrow}$ precludes the use of arbitrary unit
propagation, and so when introducing restarts, we cannot propagate using nogoods.  Instead, we can
only detect when we are inside an already-visited branch.  We must therefore use the one watched
literal scheme instead, and we also introduce a basic subsumption scheme to prune redundant clauses.

Performance results from these two modified algorithms, using the same families of instances as in
the previous section, are shown in \cref{figure:mcs}. Although we have
moved from a decision problem to an optimisation problem, the same changes remain clearly
beneficial. For the $k{\downarrow}$ algorithm, the change has a minimal effect on many instances
(typically, where the $k = 0$ subproblem is unsatisfiable and hard, and the $k = 1$ subproblem is
satisfiable and easy), but gives large benefits on many more instances than it penalises: it is over
an order of magnitude better on nearly a hundred instances, whilst being an order of magnitude worse
on only seven.

With McSplit${\downarrow}$, the inability to use two watched literals means that in many cases we
introduce a small slowdown. However, the overall pattern is the same: when introducing restarts and
a biased value ordering heuristic, it is much more common to see a large speedup than a large
slowdown.

\section{Conclusion and Future Work}

We have shown that it is both possible and beneficial to introduce a small amount of randomness into
backtracking search algorithms for hard subgraph problems, without compromising existing search
order heuristics, and without paying a penalty for unsatisfiable instances. In principle it is
well-known that introducing some form of randomness \cite{DBLP:conf/aaai/GomesSK98} or reduced
commitment \cite{DBLP:conf/ijcai/HarveyG95} into backtracking search can be beneficial, but until
now it has not been clear how to do so efficiently in practice, particularly when we already have
strong search ordering heuristics and when are expecting to encounter large numbers of unsatisfiable
instances. Our results show that the key to success is a combination of aggressive restarts, nogood
recording, and biasing an existing value-ordering heuristic with just a pinch of randomness.

?? Parallel

?? Apply to CP in general

\section*{Acknowledgements}

The authors would like to thank Ruth Hoffmann, Christophe Lecoutre, Christine Solnon, and Craig
Reilly for their comments.

\bibliographystyle{splncs04}
\bibliography{dblp}

\end{document}