### 2019-06-07 talk/talk.tex
Tweaking

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 foldmethod=marker foldmarker={{{,}}} :

\documentclass{beamer}

\usepackage{tikz}
\usepackage{xcolor}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathcal
\usepackage{amssymb}                   % \nexists
\usepackage[vlined]{algorithm2e} % algorithms
\usepackage{centernot}
\usepackage{listings}
\usepackage{csquotes}

\RequirePackage[tt=false, type1=true]{libertine}
\RequirePackage[varqu]{zi4}
\RequirePackage[libertine]{newtxmath}
\RequirePackage[T1]{fontenc}

\usetikzlibrary{shapes, arrows, shadows, calc, positioning, fit}
\usetikzlibrary{decorations.pathreplacing, decorations.pathmorphing, shapes.misc}
\usetikzlibrary{tikzmark}

\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.2, 0.129412}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
\tikzset{notvertex/.style={vertex, color=white, text=black}}
\tikzset{plainvertex/.style={vertex}}
\tikzset{vertexc1/.style={vertex, fill=uofgburgundy, text=white}}
\tikzset{vertexc2/.style={vertex, fill=uofgsandstone, text=white}}
\tikzset{vertexc3/.style={vertex, fill=uofgforest, text=white}}
\tikzset{vertexc4/.style={vertex, fill=uofgheather, text=white}}
\tikzset{edge/.style={color=black!50!white}}
\tikzset{bedge/.style={ultra thick}}

% {{{ theme things
\useoutertheme[footline=authortitle]{miniframes}
\useinnertheme{rectangles}

\setbeamerfont{block title}{size={}}
\setbeamerfont{title}{size=\large,series=\bfseries}
\setbeamerfont{section title}{size=\large,series=\mdseries}
\setbeamerfont{author}{size=\normalsize,series=\mdseries}
\setbeamercolor*{structure}{fg=uofguniversityblue}
\setbeamercolor*{palette primary}{use=structure,fg=black,bg=white}
\setbeamercolor*{palette secondary}{use=structure,fg=white,bg=uofgcobalt}
\setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=uofguniversityblue}
\setbeamercolor*{palette quaternary}{fg=white,bg=black}

\setbeamercolor*{titlelike}{parent=palette primary}

\beamertemplatenavigationsymbolsempty

\setbeamertemplate{title page}
{
    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.4cm)}] at (current page.north east) {
            \begin{minipage}{0.20\paperwidth}
                \begin{center}
                \includegraphics[keepaspectratio=true,scale=0.35]{UoG_keyline.pdf} \\[0.15cm]
                \includegraphics[keepaspectratio=true,scale=0.07]{UStA_Logo_SP.png}
                \end{center}
            \end{minipage}
        };

        \node [anchor=west, xshift=0.2cm, yshift=0.2cm] at (current page.west |- logo.west) {
            \begin{minipage}{0.65\paperwidth}\raggedright
                {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
            \end{minipage}
        };
    \end{tikzpicture}
}

\setbeamertemplate{section page}
{
    \begin{centering}
        \begin{beamercolorbox}[sep=12pt,center]{part title}
            \usebeamerfont{section title}\insertsection\par
        \end{beamercolorbox}
    \end{centering}
}

\newcommand{\frameofframes}{/}
\newcommand{\setframeofframes}[1]{\renewcommand{\frameofframes}{#1}}

\makeatletter
\setbeamertemplate{footline}
{%
    \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
    \end{beamercolorbox}
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
        \hfill%
        {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        {\usebeamerfont{title in head/foot}\insertshorttitle}%
        \hfill%
        {\usebeamerfont{frame number}\usebeamercolor[fg]{frame number}\insertframenumber~\frameofframes~\inserttotalframenumber}
    \end{beamercolorbox}%
    \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
    \end{beamercolorbox}
}

\makeatletter
\newenvironment{nearlyplainframe}[2][]{
    \def\beamer@entrycode{\vspace*{-\headheight}\vspace*{3pt}}
    \setbeamertemplate{headline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
        \end{beamercolorbox}
        \vspace*{\headheight}
    }

    \setbeamertemplate{footline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
        \end{beamercolorbox}
    }

    \begin{frame}[#1]{#2}
    }{
    \end{frame}
}
\makeatother

\makeatletter
\newenvironment{justborderframe}[2][]{
    \def\beamer@entrycode{\vspace*{-\headheight}}
    \setbeamertemplate{headline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
        \end{beamercolorbox}
        \vspace*{\headheight}
    }

    \setbeamertemplate{footline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
        \end{beamercolorbox}
    }

    \begin{frame}[#1]{}
    }{
    \end{frame}
}
\makeatother

% }}}

\author{Blair Archibald \and Fraser Dunlop \and Ruth Hoffmann \and Ciaran McCreesh \and Patrick
Prosser \and James Trimble}
\title{Sequential and Parallel Solution-Biased Search for Subgraph Algorithms}

\begin{document}

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics[keepaspectratio=true, height=\paperheight]{background.jpg}};
    }
    \begin{frame}[plain,noframenumbering]
        \titlepage
    \end{frame}
}

\begin{frame}{Subgraph Isomorphism}
    \centering
    \begin{tikzpicture}
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};

        \draw [ultra thick] (Na) -- (Nb);
        \draw [ultra thick] (Nb) -- (Nc);
        \draw [ultra thick] (Nc) -- (Nd);
        \draw [ultra thick] (Nb) -- (Nd);

        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};

        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);

        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <2> [ultra thick] (N1) -- (N3);
        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <2> [ultra thick] (N3) -- (N5);
        \draw <2> [ultra thick] (N3) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <2> [ultra thick] (N5) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
    \end{tikzpicture}
\end{frame}

\begin{frame}{Subgraph Finding, as a Constraint Program}
    \begin{itemize}
        \item A variable for each pattern vertex. The domains are all of the target vertices.
        \item At least two sets of constraints:
            \begin{itemize}
                \item Adjacent pairs of vertices must be mapped to adjacent pairs of vertices.
                \item All different.
            \end{itemize}
        \item Then we get clever:
            \begin{itemize}
                \item Extra constraints about degrees, paths, \ldots
                \item Better data structures and propagation queues.
                \item Very good variable- and value-ordering heuristics.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{The Glasgow Subgraph Solver}
    \begin{center}
        \url{https://github.com/ciaranm/glasgow-subgraph-solver}
    \end{center}

    \begin{itemize}
        \item A CP style solver specifically for subgraph algorithms.
        \item Subgraph isomorphism, and all its variants (induced / non-induced, homomorphism,
            locally injective, labels, side constraints, directed, \ldots).
        \item Also special algorithms for clique.
    \end{itemize}
\end{frame}

\begin{frame}{Benchmark Instances}
    \begin{itemize}
        \item 14,621 instances from Christine Solnon's collection:
            \begin{itemize}
                \item Randomly generated with different models.
                \item Real-world graphs.
                \item Computer vision problems.
                \item Biochemistry problems.
                \item Phase transition instances.
            \end{itemize}
        \item At least\ldots
            \begin{itemize}
                \item $\ge$ 2,110 satisfiable.
                \item $\ge$ 12,322 unsatisfiable.
            \end{itemize}
        \item A lot of them are very easy for good algorithms.
    \end{itemize}
\end{frame}

\begin{frame}{Hardware}
    \begin{itemize}
        \item HPC, optimised for throughput not reproducibility.\footnote{This work used the Cirrus
            UK National Tier-2 HPC Service at EPCC (http://www.cirrus.ac.uk) funded by the
            University of Edinburgh and EPSRC (EP/P020267/1)}
        \item Dual Intel Xeon E5-2695 v4 CPUs, $2 \times 18$ cores
        \item 256GBytes RAM
        \item GCC 7.2.0
        \item C++ native threads, SGI MPT MPI
    \end{itemize}
\end{frame}

\begin{frame}{Value-Ordering Heuristics}
    \begin{itemize}
        \item Largest target degree first.
    \end{itemize}
\end{frame}

\begin{frame}{Sanity Check}\includegraphics<1>{gen-graph-value-ordering.pdf}\includegraphics<2>{gen-graph-value-ordering-unsat.pdf}
\end{frame}

\begin{frame}{However\ldots}
    \begin{itemize}
        \item What if several vertices have the same degree?
        \item Is a vertex of degree $10$ really that much better than a vertex of degree $9$?
    \end{itemize}
\end{frame}

\begin{frame}{Discrepancy Search?}
    \begin{center}
        \includegraphics<1>{gen-graph-value-ordering-dds.pdf}%
        \includegraphics<2>{gen-graph-value-ordering-dds-unsat.pdf}%
        \includegraphics<3>{gen-graph-value-ordering-dds-scatter.pdf}
    \end{center}
\end{frame}

\begin{frame}{Value-Ordering Heuristics as Distributions}
    \begin{itemize}
        \item Traditional view: value-ordering defines a search order.
        \item New view: value-ordering defines \textcolor{uofgcobalt}{what proportion of the search
            effort} should be spent on different subproblems.
    \end{itemize}
\end{frame}

\begin{frame}{A Slightly Random Value-Ordering Heuristic}
    \only<1>{
    \begin{itemize}
        \item For a fixed domain $D_v$, pick a vertex $v'$ from a domain $D_v$ with probability
    \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \]
        \item Equally likely to pick between two vertices of degree $d$.
        \item Twice as likely to select a vertex of degree $d$ than a vertex of degree $d - 1$.
        \item In the paper: \textcolor{uofgcobalt}{solution density} and expected distribution of solutions.
    \end{itemize}}\only<2>{
    \begin{center}
        \includegraphics<2>{gen-graph-bias-scatter.pdf}%
    \end{center}}
\end{frame}

\begin{frame}{Restarts and Nogood Recording}
    \begin{itemize}
        \item Aggressive restarts: every 100ms.
        \item Nogood recording and 2WL to avoid repeating work.
    \end{itemize}
\end{frame}

\begin{frame}{Is It Better?}
    \begin{center}
        \includegraphics<1>{gen-graph-sbs.pdf}%
        \includegraphics<2>{gen-graph-sbs-unsat.pdf}%
        \includegraphics<3>{gen-graph-sbs-scatter.pdf}%
    \end{center}
\end{frame}

\begin{frame}{Parallel Search}
    \begin{itemize}
        \item Each thread gets its own random seed.
        \item Barrier synchronise on restarts.
        \item Share nogoods.
    \end{itemize}
\end{frame}

\begin{frame}{Is It Even Betterer?}
    \includegraphics<1>{gen-graph-par-scatter.pdf}%
    \includegraphics<2>{gen-graph-par.pdf}%
    \includegraphics<3>{gen-graph-dist.pdf}%
\end{frame}

\begin{frame}{Broader Perspective}
    \begin{itemize}
        \item Can scale to even more hosts with less strict synchronisation.
        \item Also good for maximum common subgraph algorithms.
        \item Does this work for CP in general?
        \item Can we finally kill off vanilla sequential backtracking search?
    \end{itemize}
\end{frame}

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner
        sep=0pt]{\includegraphics[keepaspectratio=true, width=\paperwidth]{background2.jpg}};
    }

    \begin{frame}[plain,noframenumbering]
        \begin{tikzpicture}[remember picture, overlay]
            \node at (current page.north west) {
                \begin{tikzpicture}[remember picture, overlay]
                    \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
                \end{tikzpicture}
            };

            \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
                \includegraphics[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
            };

            \node [anchor=north west, shift={(0.6cm,-0.8cm)}] at (current page.north west) {
                \textcolor{white}{\url{http://www.dcs.gla.ac.uk/~ciaran}}
            };

            \node [anchor=north west, shift={(0.6cm,-1.4cm)}] at (current page.north west) {
                    \textcolor{white}{\href{mailto:ciaran.mccreesh@glasgow.ac.uk}{\nolinkurl{ciaran.mccreesh@glasgow.ac.uk}}}
            };
        \end{tikzpicture}
    \end{frame}
}

\end{document}